### 2018-01-31 includes/introduction.tex
First pass with Federico.
Fixing up more content.

In recent years many algorithms for solving string constraints have
been developed and implemented as part of SMT solvers such as CVC4
\cite{cvc4} and Z3 (e.g., Z3str2 \cite{z3str2} and Z3str3
\cite{z3str3}). To validate and benchmark these solvers, their
developers have relied on hand-crafted input
suites~\cite{cvc4-tests,z3str3-tests,z3str2-tests} or real-world
examples from a limited class of industrial
applications~\cite{kaluza,kausler}. These test suites have helped
developers identify implementation defects and develop more
sophisticated solving heuristics. Although extremely helpful, these
suites are of limited value since they do not test the vast majority
of interesting behaviors in solvers. There is an acute need for a more
robust, inexpensive, and automatic way of generating benchmarks to test
the correctness and performance of SMT solvers, that can augment
hand-crafted as well as tests generated from industrial applications.

%VG: add citation to SAT fuzzer
Random and mutation fuzzing are widely used to test all kinds of software
including SAT solvers. Inspired by their utility, we introduce a
fuzzer for string solvers, \fuzzer{}, and describe its value as an
exploratory validation tool. We demonstrate the efficacy of \fuzzer{}
by presenting defects and limitations it enabled us to identify in
leading string solvers, which otherwise would have been difficult to
detect. To the best of our knowledge, \fuzzer{} is the first such tool
aimed at automatic generation of string constraints. \fuzzer{} can be
used to mutate or transform existing benchmarks, as well as
randomly generate structured instances that can be scaled with respect
to a variety of parameters. Examples include: length of string constants,
depth of concats, balanced or unbalanced concats, overlapping word equations,
systems of word equations with arithmetic over length, nested regular
expressions, rich combinations of regular expressions and string
constraints, and many more.

\noindent{\textbf{Contributions:}}

\begin{enumerate}
\item \textbf{The \fuzzer{} tool}: We describe in
  Section~\ref{sec:fuzzer} a modular fuzzer which can transform and
  randomly generate extremal \smtfull{} string and regular expression
  (regex) problems~\footnote{We assume basic
    familiarity with string solvers and their input
    language.}. Extremal or scaling inputs (e.g., long string
  constants, deep concatenations) are particularly useful in
  identifying asymptotic behaviors in solvers. We briefly document \fuzzer{}'s
  components and modular architecture. We provide example use cases to
  demonstrate its utility as an exploratory tool for validating
  solvers.

\item \textbf{A repository of \smtfull{} problems}: We present a
  repository of \smtfull{} string and regex problem suites we
  generated using \fuzzer{} in Section~\ref{sec:problems}. We
  demonstrate the usefulness of these suites by revealing specific
  defects they helped uncover and fix in \us{}.

\item \textbf{Experimental Results}: We compare \theSolvers{} on the
  \fuzzer{} suites and selectively present the experimental result on
  the suites \theSuites{} in Section~\ref{sec:data}. We highlighted
  these suites because they uniquely made some solvers perform poorly,
  but not others.

\item \textbf{Analysis}: We analyze our experimental results in
  Section~\ref{sec:analysis} by inspecting the execution traces of the
  solvers on the given problems. We pinpoint algorithmic limitations
  in \us{} that cause poor performance, and analyze proposed and
  implemented enhancements that address the issues which \fuzzer{}
  helped us to identify.
\end{enumerate}

>>>>

In recent years many algorithms for solving string constraints have
been developed and implemented in SMT solvers such as Norn~\cite{norn},
CVC4~\cite{cvc4}, and Z3 (e.g., Z3str2~\cite{z3str2} and Z3str3~\cite{z3str3}).
To validate and benchmark these solvers, their developers have relied on
hand-crafted input suites~\cite{cvc4-tests,z3str3-tests,z3str2-tests} or
real-world examples from a limited class of industrial
applications~\cite{kaluza,kausler}. These test suites have helped
developers identify implementation defects and develop more
sophisticated solving heuristics. Unfortunately, as solvers grow,
fixed benchmarks leave more functionality untested.
There is an acute need for a more robust, inexpensive, and automatic way
of generating benchmarks to test
the correctness and performance of SMT solvers.

%VG: add citation to SAT fuzzer
Fuzzing is often used to test all kinds of software,
including SAT solvers. Inspired by their utility, we introduce a
fuzzer for string solvers, \fuzzer{}, and describe its value as an
exploratory validation tool. We demonstrate its efficacy
by presenting limitations it helped us find in
leading string solvers. To the best of our knowledge, \fuzzer{} is the
first tool aimed at automatic generation of string constraints. \fuzzer{} can be
used to mutate or transform existing benchmarks, as well as
randomly generate structured instances. These instances can be scaled with respect
to a variety of parameters, e.g., length of string constants,
depth of concats and regeular expressions, number of variables,
number of length constraints, and many more.

\subsubsection{Contributions}

\begin{enumerate}
\item \textbf{The \fuzzer{} tool}:
    In section~\ref{sec:fuzzer}, we describe a modular fuzzer that can transform and
    randomly generate extremal \smtfull{} string and regular expression
    (regex) problems~\footnote{We assume basic
    familiarity with string solvers and their input
    language.}. Extremal or scaling inputs (e.g., long string
    constants, deep concatenations) are particularly useful in
    identifying asymptotic behaviors in solvers. We briefly document \fuzzer{}'s
    components and modular architecture. We provide example use cases to
    demonstrate its utility as an exploratory tool for validating
    solvers.

\item \textbf{A repository of \smtfull{} problems}:
    We present a repository of \smtfull{} string and regex problem suites we
    generated using \fuzzer{} in Section~\ref{sec:problems}. The repository
    consists of two categories: \texttt{generated}, with new problems generated
    by \fuzzer{}, and \texttt{transformed}, with transformed problems generated
    from a small suite of industrial benchmarks.

\item \textbf{Experimental Results and Analysis}:
    We compare the performance of \theSolvers{} on the
    the \fuzzer{} suites \theSuites{} in section~\ref{sec:data}. We highlight
    these suites because they make some solvers perform poorly,
    but not others. In section~\ref{sec:analysis}, we analyze our
    experimental results, and pinpoint algorithmic limitations
    in \us{} that cause poor performance. We propose enhancements that
    address these issues, which \fuzzer{} helped us find.
\end{enumerate}

### 2018-01-31 includes/problems.tex
Fixing up more content.

We generated several suites
of string and regex problems using \fuzzer{}\footnote{For review
anonymity, we removed the link to the source code and the generated
suites. They will be provided separately in the artifact
submission.}. Table~\ref{tbl:generated} describes the suite
\texttt{generated} by \generator{}.
\begin{table}[H]
    \centering
    \caption{\generator{}-generated problem sets.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}             & \textbf{Problems have ...} \\ \hline
        \textit{Concats}          & A single, deep tree of concats. \\ \hline
        \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
        \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
        \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
        \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
        \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
        \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
        \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
        \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
        \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
\end{table}

>>>>

We generated several suites
of string and regex problems using \fuzzer{}\footnote{The link to the
repository will be added after double-blind review.}.
Table~\ref{tbl:generated} describes the suite
\texttt{generated} by \generator{}.
\begin{table}[H]
    \centering
    \caption{\generator{}-generated problem sets.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}             & \textbf{Problems have ...} \\ \hline
        \textit{Concats}          & A single, deep tree of concats. \\ \hline
        \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
        \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
        \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
        \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
        \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
        \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
        \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
        \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
        \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
\end{table}

### 2018-01-31 includes/stringfuzz.tex
Fixing up more content.

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string
problems. It can either be installed from source, or from the Python
PIP package repository. \fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. Specifically, \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} problems. It supports several generators and
        options that specify its output. Details can be found in
        Table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        problems. It supports several transformers and options that specify
        its output and input, which are explained in
        Table~\ref{tbl:transformers}. Note that transformers
        \textit{translate} and \textit{reverse} also preserve problem
        satisfiability \cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        problem as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[H]
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
        \textit{concats}     & Long concats and optional random extracts. \\ \hline
        \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{regex}       & Regexes of varying complexity. \\ \hline
        \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
\end{table}

\begin{table}[H]
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{nop}         & Does nothing. \\ \hline
        \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

@@

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms. Each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are several example regexes
(with spacing added for clarity) of depth 2 produced by this scheme:
\begin{align*}
  & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad
  ((\texttt{ddd})*)+\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}

>>>>

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
problems. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} problems. It supports several generators and
        options that specify its output. Details can be found in
        Table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        problems. It supports several transformers and options that specify
        its output and input, which are explained in
        Table~\ref{tbl:transformers}. Note that transformers
        \textit{translate} and \textit{reverse} also preserve problem
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        problem as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[H]
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
        \textit{concats}     & Long concats and optional random extracts. \\ \hline
        \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{regex}       & Regexes of varying complexity. \\ \hline
        \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{nop}         & Does nothing. \\ \hline
        \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

@@

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms. Each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are three example regexes
(separated by spaces) of depth 2 produced by this scheme:
\begin{align*}
  & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad\quad
  ((\texttt{ddd})*)+\quad\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}
